---
title: "Homework #1"
output: 
  html_document: 
    df_print: kable
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

## Data and packages

```{r}
## install tidyverse if not alrady
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

## load tidyverse
library(tidyverse)

## read-in functions for reading/wrangling qualtrics data
source("https://github.com/mkearney/stat/raw/master/static/R/read_qualtrics.R")

## read in qualtrics data
d <- suppressMessages(
  readr::read_csv("https://github.com/mkearney/stat/raw/master/data/elect16data2.csv"))
```

## 1. How many observations (rows)?

```{r}
## your code goes in here. it should output the correct number of rows
nrow(d)
```


## 2. How many variables (columns)?
```{r}
## your code goes in here. it should output the correct number of variables=
ncol(d)
```


## 3. What are all of the possible responses for `ambiv_sexism_7`

```{r}
## your code goes in here. it should output the a frequency table for the correct variable
table(d$ambiv_sexism_7)
```


## 4. Recode all likert-type columns

```{r}
## this one is done for you (this function was sourced in earlier)
d <- recode_likerts(d)
```

## 5. Calculate the mean of `ambiv_sexism` row means

```{r}
## your code goes in here. first, calculate means (for ambiv_sexism items) for each row
d$ambiv_sexism <- rowMeans(dplyr::select(d, ambiv_sexism_1:ambiv_sexism_22), na.rm = TRUE)

## second, calculate the mean from the totals
mean(d$ambiv_sexism, na.rm = TRUE)
```

## 6. Calculate the standard deviation of `ambiv_sexism` means

```{r}
## hint: use the means you calculated in the previous question
sd(d$ambiv_sexism, na.rm = TRUE)
```

## 7. How many respondents were there in each of the observed race categories?

```{r}
## your code goes in here. it should output a frequency table of the race variable
table(d$race)
```

## 8. What is the median age of the respondents?

```{r}
## your code goes in here. it should output the median of the age variable
median(d$age, na.rm = TRUE)
```

## 9. Use psych::describe to summarize `ambiv_sexism`, `therm_1`, `therm_2`, `therm_3`, `therm_4`, `therm_5`

```{r}
## your code goes in here. it should output the summary data from pscyh::describe
psych::describe(d[, c("ambiv_sexism", "therm_1", "therm_2", "therm_3", "therm_4", "therm_5")])
```

## 10. Plot histograms of `therm_1`, `therm_2`, `therm_3`, `therm_4`

```{r}
## histograms with base R
hist(d$therm_1, col = "#3366ff")
hist(d$therm_2, col = "#dd3333")
hist(d$therm_3, col = "gray")
hist(d$therm_4, col = "hotpink")

## histograms with ggplot2
d %>%
  select(therm_1:therm_4) %>%
  gather(var, val) %>%
  ggplot(aes(x = val, fill = var)) +
  geom_histogram(bins = 20, na.rm = TRUE) +
  facet_wrap(~ var, scales = "free") + 
  theme_bw() + 
  theme(legend.position = "none", 
        plot.title = element_text(face = "bold"), 
        axis.title = element_text(hjust = .95, face = "italic")) + 
  labs(title = "Feeling thermometer distributions", 
       subtitle = "Histograms of responses (N = 254) to political feeling thermometers", 
       x = "Temperature", y = "Observations")
```