---
title: "Homework #1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

```{r}
## read-in functions for reading/wrangling qualtrics data
source("https://github.com/mkearney/stat/raw/master/static/slides/R/read_qualtrics.R")

## read in qualtrics data
d <- read_qualtrics("https://github.com/mkearney/stat/raw/master/static/slides/data/elect16data.csv")
```

## 1. How many observations (rows)?

```{r}
## your code goes in here. it should output the correct number of rows

```


## 2. How many variables (columns)?
```{r}
## your code goes in here. it should output the correct number of variables=

```


## 3. What are all of the possible responses for `ambiv_sexism_7`

```{r}
## your code goes in here. it should output the a frequency table for the correct variable

```


## 4. Recode all likert-type columns

```{r}
## this one is done for you (this function was sourced in earlier)
d <- recode_likerts(d)
```

## 5. Calculate the mean of `ambiv_sexism` totals

```{r}
## your code goes in here. first, add up totals (for ambiv_sexism items) for each row

## second, calculate the mean from the totals

```

## 6. Calculate the standard deviation of `ambiv_sexism` totals

```{r}
## hint: use the sums you calculated in the previous question

```

## 7. How many respondents were there in each of the observed race categories?

```{r}
## your code goes in here. it should output a frequency table of the race variable

```

## 8. What is the median age of the respondents?

```{r}
## your code goes in here. it should output the median of the age variable

```

## 9. Use psych::describe to summarize the entire data set

```{r}
## your code goes in here. it should output the summary data from pscyh::describe

```

## 10. Plot histograms of `therm_1` and `therm_2`

```{r}
## your code goes in here. it should output 2 histogram plots

```