---
title: "Homework #2"
output: html_notebook
---

The data set `papp`, entered below, provides President approval ratings (on 0-10 scale) given by an equal number of journalists and non-journalists.

```{r}
## pres approval rating data
papp <- data.frame(
  non_journo = c(3, 2, 5, 7, 1, 2, 5, 4, 3, 9, 2),
  journalist = c(0, 2, 1, 4, 2, 6, 1, 3, 2, 4, 2)
)
```

1. Calculate (and store) the mean `x1` `x2`, variance `s1` `s2`, and size `n1` `n2` of each group.

```{r}
## your code goes here
```

2. Using the objects created for the previous question, calculate (and store as `num` and `denom`) the numerator and denominator based on the independent samples t-test formula

```{r}
## your code goes here
```

3. Using `num` and `denom`, calculate (and store as `tstat`) the t statistic

```{r}
## your code goes here
```

4. Calculate (and store as `df`) the degrees of freedom (you can assume equal variances)

```{r}
## your code goes here
```

5. Use `tstat` and `df` to calculate the p-value for a two-sided hypothesis

```{r}
## your code goes here
```

6. Use `tstat` and `df` to calculate the p-value for a one-sided (non-journalist greater than journalist) hypothesis

```{r}
## your code goes here
```

7. Using the `papp` data set, use the `t.test()` function to conduct a t-test, comparing the approval ratings between the two groups.

```{r}
## your code goes here
```

8. Based on the p-values you calculated for the last two questions, is either of the tests significant? Explain why or why not for each.

<!-- type your answer below here; you don't need to provide R code for this one -->

9. Visualize the means or distributions of the two groups. Make sure the values associated with each group are clearly marked.

```{r}
## your code goes here
```

