---
title: "Homework #3"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

The data set `d`, entered below, provides estimates on media use, health info, and education levels.

```{r}
## pres approval rating data
d <- data.frame(
  media_use = c(3, 2, 5, 7, 1, 2, 5, 4, 3, 9, 2),
  health_info = c(0, 2, 1, 4, 2, 6, 1, 3, 2, 4, 2),
  edu = c(3, 2, 4, 2, 1, 5, 2, 3, 4, 5, 3)
)
```

1. Calculate Pearson's correlation coefficient between media use and health information.

```{r}
## your code goes here
cor(d$media_use, d$health_info, method = "pearson")
```

2. Calculate the rank-order correlation coefficient between media use and health information. 

```{r}
## your code goes here
cor(d$media_use, d$health_info, method = "spearman")
```

3. Conduct a signifance test of the correlation between media use and health information. Use the appropriate method.

```{r}
## your code goes here
cor.test(d$media_use, d$health_info)
```

4. Create a correlation matrix containing all three variables

```{r}
## your code goes here

```

5. Visualize the correlations

```{r}
## your code goes here

```

