---
title: "R Notebook"
output: html_notebook
---

## the data


```{r}
if (!requireNamespace("broom", quietly = TRUE)) {
  install.packages("broom")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
## load tidyverse
library(tidyverse)

## read in exam 1 data
examdat <- readr::read_csv("https://github.com/mkearney/stat/raw/master/static/data/1-data.csv", skip = 1)

## create news_use, pie, and cat variables
examdat <- examdat %>%
  mutate(news_use = tv + news,
    pie = (pie1 + pie2 + pie3 + pie4) / 4,
    cat = sample(letters[1:4], nrow(examdat), replace = TRUE))
```


## 1. Simple (one variable) models

Linear model for `news_use` predicted by `sex`

```{r}
m1a <- lm(news_use ~ 1 + sex, data = examdat)
summary(m1a)
```


```{r}
m1b <- lm(news_use ~ 0 + sex, data = examdat)
summary(m1b)
```


## 2. Simple (one variable) multiple-level models

Linear model for `news_use` predicted by `race`

```{r}
m2a <- lm(news_use ~ 1 + race, data = examdat)
summary(m2a)
```


```{r}
m2b <- lm(news_use ~ 0 + race, data = examdat)
summary(m2b)
```





## 3. Two-way ANOVA

Linear model for `news_use` predicted by `race` and `sex`

```{r}
m3a <- lm(news_use ~ 1 + race + sex, data = examdat)
summary(m3a)
```


```{r}
m3b <- lm(news_use ~ 0 + race + sex, data = examdat)
summary(m3b)
```




## 4. Interaction

Linear model for `news_use` predicted by `race` and `sex` interaction

```{r}
m4a <- lm(news_use ~ 1 + race * sex, data = examdat)
summary(m4a)
```


```{r}
m4b <- lm(news_use ~ 0 + race * sex, data = examdat)
summary(m4b)
```


## 5. Compare models

Compare the models with the intercept 

```{r}
anova(m3a, m4a)
```

Compare the models without the intercept

```{r}
anova(m3b, m4b)
```
