---
title: "Visualize Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load the [tidyverse](https://tidyverse.org) set of packages.

```{r setup}
## load tidyverse
library(tidyverse)
```

The data set:

```{r}
## preview mpg data
mpg
```

## Quiz

What relationship do you expect to see between engine size (`displ`) and highway fuel efficiency (`hwy`)?

## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}


```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = hwy, 
    color = model, 
    alpha = displ, shape = year > 2000))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
mpg %>%
  mutate(class = case_when(
    class == "2seater" ~ "Two-Seater",
    class == "compact" ~ "Compact",
    class == "subcompact" ~ "Sub-Compact",
    class == "midsize" ~ "Mid-Size",
    class == "minivan" ~ "Mini-Van",
    class == "pickup" ~ "Pickup Truck",
    class == "suv" ~ "S.U.V.", 
    TRUE ~ NA_character_
  )) %>%
  mutate(class = factor(class, levels = c("Sub-Compact","Two-Seater", "Compact", 
    "Mid-Size", "PickupTruck", "Mini-Van", "S.U.V."))) %>%
  ggplot() + 
  geom_boxplot(aes(x = class, y = hwy, fill = class, color = class), alpha = .5) + 
  geom_jitter(aes(x = class, y = hwy, fill = class), shape = 21, size = 2.5, alpha = .5) + 
  theme_minimal() + 
  labs(title = "Highway fuel efficiency by car class", 
    subtitle = "Based on data from the ggplot2::mpg data set", 
    x = "Class of Car", y = "Highway Fuel Efficiency (MPG)") + 
  theme(plot.title = element_text(face = "bold"), 
    axis.title = element_text(face = "italic", hjust = .95))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}
ggplot(data = mpg) + geom_histogram(aes(x = hwy), binwidth = 2)

```

## Your Turn 5

Make a density plot of `hwy` colored by `class`.

```{r}
ggplot(data = mpg) + geom_density(aes(x = hwy), fill = "red", 
  alpha = .5, binwidth = 2, bw = 1) + 
  geom_histogram(aes(x = hwy, y = ..density..), alpha = .25)
```

## Your Turn 6

Make a bar chart `class` colored by `class`.

```{r}
ggplot(data = mpg) + 
  geom_bar(aes(x = class, fill = class), width = .2) + 
  coord_flip()
```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy), method = "lm")
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy), method = "lm") + 
  ggsave("displ-hw.png", width = 8, height = 5, units = "in")
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```
