---
title: 'Exam #2'
author: 'YOUR NAME GOES HERE'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exam #2

The data for this exam come from Amazon mturk workers who were paid 
35 cents to copmlete a Qualtrics questionnaire. In addition to basic
demographic items, respondents were asked about their political
orientations and media habits. Also featured in the questionnaire was a
2 (headline formats: twitter.com/apnews.com) x 2 (headline topics: 
politics/health) experiment wherein each respondent was randomly
exposed to three news headlines that were either about health or politics 
and either formatted as though they came from apnews.com or twitter.com.

The names of the variables in the data set are fairly straightforward, but 
you should also still have access to the Qualtrics survey, so you'll probably
want to spend a little bit of time making sure you understand which Qualtrics
item produced which column in the data set.

To read the data set, use the following code:

```{r}
## read in the twitter-AP news headline study data
twap <- readRDS(
  "https://github.com/mkearney/stat/blob/master/static/exams/exam2-data.rds?raw=true"
)
```

## 1. Descriptives

For your first task, I'd like you to describe the data set. Be sure to include
any code you use in the code chunk below, but then write a paragraph like you 
would in a full-on research paper. At a minimum, the paragraph should include
the following details:

+ The number of respondents (sample size)
+ The number and percent of each `gender`
+ The number and percent of each `race`
+ The number and percent of each political `party`
+ The mean, standard deviation, minimum, and maximum of `age`
+ The modal (mode) response to the education question (you'll need to look up
the corresponding response label from Qualtrics)

```{r}
## descriptives-related code goes here


```

<!-- ------------------------------------------------- -->
<!-- 1. DESCRIPTIVES PARAGRAPH WRITE-UP SHOULD GO HERE -->
<!-- ------------------------------------------------- -->

## 2-3. Hypotheses

With this study, we're ultimately interested in whether the presentation
(`headline_format`) and/or topic (`headline_topic`) of news headlines have 
an effect on the perceived newsworthiness (`newsworthy`) and/or credibility 
(`credibility_*`) of the information. With this in mind, write two 
hypotheses–one predicting each DV (`newsworthy` and `credibility`). Each
hypothesis should theorize a relationship between `headline_format` and 
`headline_topic` and the dependent variable. Whether you want to theorize
about an interaction between the two variables is up to you.

<!-- ------------------------------------------------- -->
<!-- H1 about `newsworthy` goes about here             -->
<!-- ------------------------------------------------- -->

<!-- ------------------------------------------------- -->
<!-- H2 about `credibility` goes about here            -->
<!-- ------------------------------------------------- -->

## 3-5. Credibility

As you probably noticed, we have seven different `credibility_*` 
variables–each of which represents responses to bipolar semantic
scales with higher numbers associated with the more positive of 
the two words and lower numbers associated with the more
negative of the two words. For your third and fourth tasks, I want 
you to decide which of those seven items you'd like to use to create
your credibility variable (and then create/add a `credibility` 
variable to the data set) and then explain your decision below. And, 
for the record, there's not really a right or wrong answer, but 
whatever you decide to do should be supported by both *theoretical* 
(logic, reason, human understanding of media/communication/journalism, 
etc.) and *empirical* reasons (correlations, factor analysis, etc.).

```{r}
## insert code below that adds a `credibility` variable to the data set
## the two lines below are two possible examples of what the code may 
## start to look like (only pick one to work from)
#twap$credibility <- 
#twap <- mutate(twap, credibility = )


```

<!-- ------------------------------------------------- -->
<!-- Paragraph (5-6 sentences) about credibility here  -->
<!-- ------------------------------------------------- -->

## 6. Hypothesis #1

Your first hypothesis should predict relationship(s) between the format
and topic of headlines and the perceived newsworthiness of the headline.
To test your first hypothesis, I'd like you to use Analysis of Variance 
(ANOVA) via the `aov()` function. In the code chunk below, estimate an 
`aov()` model, print out a summary of the results, and then write a 
brief description of what you found--if your results are significant, be 
sure to point out which level of theexperimental factor(s) corresponded 
with higher and lower levels of newsworthiness.

*NOTE: Make sure you formally state whether the results reject or fail 
to reject the NULL*

```{r}
## anova model code goes here

```

<!-- ------------------------------------------------- -->
<!-- Write-up of H1 results goes about here            -->
<!-- ------------------------------------------------- -->


## 7. Hypothesis #2

For your second hypothesis, which should predict relationship(s) between 
the format and topic of headlines and the credibility of the source of
the headline, I want you to use regression via the `lm()` function and include 
**at least four control variables** (other variables to be included as
predictors besides `headline_format` and `headline_topic`) in the model.
In the code chunk below you should again provide code to estimate the 
model. And then in the space below that, provide a brief summary of the
results and a brief explanation of why you chose the control variables 
that you did.

*NOTE: Make sure you formally state whether the results reject or fail 
to reject the NULL*

```{r}
## regression model code goes here


```

<!-- ------------------------------------------------- -->
<!-- Write-up of H2 results goes about here            -->
<!-- ------------------------------------------------- -->

## 8. #dataviz

For your final task, I'd like you create a data visualization for your 
first hypothesis. In one way or another, your data visualization should 
compare the means and/or distributions of the four groups (twitter 
versus apnews; health versus politics).

```{r}
## you data viz code and output go below here
```

