---
title: 'Exam #2'
author: 'YOUR NAME GOES HERE'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Exam #2

Data for this exam were collected from mturk workers via Qualtrics. To allow for
more of a focus on the statistical techniques, I've done a lot of the data cleaning
for you. I think the variable names are fairly straightforward, but even if they
aren't, you should still have access to the Qualtrics survey, so you're encouraged
to spend a little bit of time becoming familiar with the survey–e.g., making sure
you have a decent understanding of which columns were produced by which Qualtrics
items. I've also included a write-up of the data collection procedures (methods),
so hopefully that helps provide some context for the data as well. 

## Method

### Procedures

Data collection consisted of an online questionnaire, which was distributed to 
United States-based Amazon Mturk workers who were paid 35 cents to complete a 
Qualtrics questionnaire "about their opinions on the news." In addition to basic
demographic items, respondents were asked a small handful of questions about their 
political orientations and media habits. The questionnaire then featured stimuli 
for a 2 (headline formats: twitter.com/apnews.com) x 2 (headline topics: 
politics/health) factorial-design experiment. All respondents were randomly
exposed to three news headlines with (a) roughly half of the respondents receiving
headlines about health while the other half received headlines about politics and 
(b) roughly half of the respondents receiving headlines formatted as tweets on 
*twitter.com* while the other half received headlines formatted as news stories 
on *apnews.com*.

### Data

To read the data set into R, use the following code:

```{r}
## download data set to temporary file
download.file(
  "https://github.com/mkearney/stat/blob/master/static/exams/exam2-data.rds?raw=true",
  assign("tmp", tempfile(fileext = ".rds"))
)
## read in the twitter-AP news headline study data
twap <- readRDS(tmp)
```

#### 1. Descriptives

For your first task, you'll need to describe the data set. Be sure to include
any code you use in the code chunk below, but then write a paragraph like you 
would in a full-on research paper. At a minimum, the paragraph should include
the following details:

+ The number of respondents (sample size)
+ The number and percent of each `gender`
+ The number and percent of each `race`
+ The number and percent of each political `party`
+ The mean, standard deviation, minimum, and maximum of `age`
+ The modal (mode) response to the education question (you'll need to look up
the corresponding response label from Qualtrics)
+ The number and percent of respondents who passed the manipulation check `mancheck`

```{r}
## descriptives-related code goes here


```

<!-- ------------------------------------------------- -->
<!-- 1. DESCRIPTIVES PARAGRAPH WRITE-UP SHOULD GO HERE -->
<!-- ------------------------------------------------- -->

## 2-3. Hypotheses

With this study, we're ultimately interested in whether the presentation
(`headline_format`) and/or topic (`headline_topic`) of news headlines have 
an effect on the perceived newsworthiness (`newsworthy`) and/or credibility 
(`credibility_*`) of the information. With this in mind, write two 
hypotheses–one predicting each DV (`newsworthy` and `credibility`). Each
hypothesis should theorize a relationship between `headline_format` and 
`headline_topic` and the dependent variable. Whether you want to theorize
about an interaction between the two variables is up to you.

<!-- ------------------------------------------------- -->
<!-- H1 about `newsworthy` goes about here             -->
<!-- ------------------------------------------------- -->

<!-- ------------------------------------------------- -->
<!-- H2 about `credibility` goes about here            -->
<!-- ------------------------------------------------- -->

## 3-4. Credibility

As you probably noticed, we have seven different `credibility_*` variables–each
one represents responses to bipolar semantic scales with higher numbers
associated with the more positive of the two words and lower numbers associated
with the more negative of the two words. For your third and fourth tasks, decide
which among the seven credibility items you think best represent the concept of
credibility (you can pick anywhere from 2-7 items) and then aggregate those
items into a single `credibility` variable (which should be added as a column to
the `twap` data object) and then explain your decision below. For the record,
there's not really a right or wrong answer, but whatever you decide to do should
be supported by both *theoretical* (logic, reason, human understanding of
media/communication/journalism, etc.) and *empirical* reasons (correlations,
factor analysis, etc.).

```{r}
## insert code below that adds a `credibility` variable to the data set
## the two lines below are two possible examples of what the code may 
## start to look like (only pick one to work from)
#twap$credibility <- 
#twap <- mutate(twap, credibility = )


```


<!-- ------------------------------------------------- -->
<!-- 5-6 sentences justifying credibility items here   -->
<!-- ------------------------------------------------- -->

## 5. Hypothesis #1

Your first hypothesis should predict relationship(s) between the format
and topic of headlines and the perceived newsworthiness of the headline.
To test your first hypothesis, use an Analysis of Variance 
(ANOVA) via the `aov()` function. In the code chunk below, estimate an 
`aov()` model, print out a summary of the results, and then write a 
brief description of what you found--if your results are significant, be 
sure to point out which level of the experimental factor(s) corresponded 
with higher and lower levels of newsworthiness.

*NOTE: Make sure you formally state whether the results reject or fail 
to reject the NULL*

```{r}
## anova model code goes here

```

<!-- ------------------------------------------------- -->
<!-- Write-up of H1 results goes about here            -->
<!-- ------------------------------------------------- -->


## 6. Hypothesis #2

For your second hypothesis, which should predict relationship(s) between 
the format and topic of headlines and the credibility of the source of
the headline, I want you to use regression via the `lm()` function and include 
**at least four control variables** (other variables to be included as
predictors besides `headline_format` and `headline_topic`) in the model.
In the code chunk below you should again provide code to estimate the 
model. And then in the space below that, provide a brief summary of the
results and a brief explanation of why you chose the control variables 
that you did.

*NOTE: Make sure you formally state whether the results reject or fail 
to reject the NULL*

```{r}
## regression model code goes here


```

<!-- ------------------------------------------------- -->
<!-- Write-up of H2 results goes about here            -->
<!-- ------------------------------------------------- -->

## 7. #dataviz

For your final task, create a data visualization for your 
first hypothesis. In one way or another, your data visualization should 
compare the means and/or distributions of the four groups (twitter 
versus apnews; health versus politics).

```{r}
## you data viz code and output go below here
```

