---
title: "Day 3"
subtitle: "Hypothesis tests and group comparisons"
author: '`r rmdees::title_author()`'
date: '`r rmdees::title_icons()`'
output:
  xaringan::moon_reader:
    lib_dir: lib
    css: ["robot", "robot-fonts", "css/slides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = "#>")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

## Agenda

---

## Agenda
1. Hypothesis testing
   - Alternative hypothesis
   - Null hypothesis
   - Model comparisons
1. Simple group comparisons
   - t-tests
   - chi-square test

---
class: inverse, center, middle

## Hypothesis testing

---

## Alternative hypothesis
+ Also referred to as **research hypothesis**, this hypothesis formally states a **theorized relationship or difference**
   - Relationship between two or more **variables**
   - Difference between two or more **groups**
+ Research hypotheses describe **patterns**, not a lack of patterns
   - Every research hypothesis has a null hypothesis
+ Alternative/research hypotheses are typically represented as $H_{1}$:...


---

## Null hypothesis
+ The **null hypothesis** is the version of the research hypothesis that assumes **no relationship or difference**
   - **No relationship** between two or more variables
   - **No difference** between two or more groups
+ Null hypotheses are typically represented as $H_{0}$:...

> In other words, a null hypothesis restates the research hypothesis as though everything in the world **only varies at random**

---

## Hypothesis example #1
+ **Research hypothesis**:
   - $H_{1}$: Journalists subscribe to more newspapers than do non-journalists.
+ **Null hypothesis**:
   - $H_{0}$: Journalists subscribe to the same number of newspapers as do non-journalists.

---

## Hypothesis example #2
+ **Research hypothesis**:
   - $H_{1}$: Exposure to health information is positively related to health knowledge.
+ **Null hypothesis**:
   - $H_{0}$: Exposure to health information is not related to health knowledge.


---

## NHST
**Null Hypothesis Significance Testing** (NHST) is the leading research paradigm

1. Theorize a relationship or difference ( $H_{1}$ )
1. Operationalize study with testable hypothesis ( $H_{0}$ )
1. Collect data
1. Conduct statistical test of $H_{0}$
1. Report whether test rejected or failed to reject $H_{0}$
1. Discuss findings as evidence (or a lack of evidence) supporting $H_{1}$

---

## Testing the null
+ We may ultimately be interested in a difference or relationship, but **we technically test the absence** of a difference or relationship
+ Positive (or significant) results **do not prove** the research hypothesis; they **only provisionally**...
   - disprove the null (i.e., *reject the null*)
   - provide evidence in support of the research hypothesis

> Scientific knowledge accumulates over time and through multiple iterations

---

## Model comparison
+ An alternative to the NHST is **model comparisons**
+ Instead of comparing a model (theorized relationship or difference) to the null, this approach compares one model to another model
   - The other model could be the null model
   - The other model could also be a simplified or theoretically-driven model
+ Use cases
   - When the null model isn't realistic
   - When comparing two or more theories


---
class: inverse, center, middle

## Group comparisons

---

## Comparing two groups

A hypothesis can theorize about...

+ relationship between two ordinal/interval/ratio-level **variables** (next week)
+ difference in one variable between two **groups** (today)

---

## Group comparisons

+ [Wilcoxon-Mann-Whitney or t-test](https://www.ncbi.nlm.nih.gov/pubmed/20414472)
+ [Unbalanced groups Wilcoxon-Mann-Whitney test](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4792850/)
+ Online tool: http://www.ccb.uni-saarland.de/software/wtest/
+ [t-tests with small Ns](https://pareonline.net/pdf/v18n10.pdf)
+ [Five-point likert items t-test vs Wilcoxon-Mann-Whitney](https://www.pareonline.net/getvn.asp?v=15&n=11)
+ [Unequal variance t-test is underused alternative to Student's t-test and WMW test](https://academic.oup.com/beheco/article/17/4/688/215960)



---
class: inverse, middle, center

## t-tests

---

## t-test
- *t-tests* are statistical techniques that test for differences between two groups
- There are multiple types of t-tests:
  - One sample
  - Paired sample
  - Independent sample

---

## t-test assumptions
- All t-tests assume the dependent variable, or the variable on which
  groups may differ, is numeric (interval or ratio)
- All t-tests assume the independent variable, or grouping variable,
  is categorical (nominal or dicthomous)


---

## One sample t-test
- Compare sample mean versus population mean using a **one sample
  t-test**
$$H_{1}: \mu \neq \bar{x}$$


---

## Paired sample t-test
- Compare the means of one group at different times using **paired
    sample t-test**
$$H_{1}: \bar{x_{t1}} \neq \bar{x_{t2}}$$


---

## Independent sample t-test
  - ***Most common***: Compare the means of two [independent] groups
    using an **independent samples t-test**
$$H_{1}: \bar{x_{1}} \neq \bar{x_{2}}$$


---

## Independent samples t-test
- To calculate an independent samples t statistic

$$t = \frac{\bar{x}_{1} - \bar{x}_{2}}{\sqrt{
\frac{s_{1}^2}{n_{1}} +
\frac{s_{2}^2}{n_{2}}}}$$


---

## t-test example
- Say we conduct an experiment with two groups and we're interested in
what causes people to actually read news articles.

- One group is given a 30 min lecture about the importance of civic
  awareness.
- The other group is told they will be debating one another on topics
  related to current events.
- Each person in both groups is given a newspaper. And then
  researchers record how much each person read.

---

## t-test example
- Our research hypothesis is that competitive goals are better drivers
  of news use than informed citizen values.
$$H_{1}: x_{debate} > x_{informed}$$

- The null hypothesis assumes there is no difference.
$$H_{0}: x_{debate} = x_{informed}$$


---

## t-test
- The output of a t-test is the t- statistic.
- Once the t statistic is computed, you can find the corresponding p
  value using the appropriate degrees of freedom and a t-table.

---

## Example t-test

A researcher wants to find out if people are happier reading about local or national human interest stories. They have 6 people read a local human interest story, and 6 different people read a national human interest story and then ask all 12 people to rate how happy they are on a scale of 1-10, with 10 being the highest rating possible.

---

## Hypothesis: two-tail

**Research hypothesis**: people reading local human interest stories experience different levels of happiness compared to people reading national human interest stories.

$$H_{1}:~~happy_{~local }~\neq~happy_{~national}$$

**Null hypothesis**: people reading local human interest stories do **not** experience different levels of happiness compared to people reading national human interest stories.

$$H_{0}:~~happy_{~local }~=~happy_{~national}$$


---

## Hypothesis: one-tail

**Research hypothesis**: people reading local human interest stories are happier than people reading national human interest stories.

$$H_{1}:~~happy_{~local }~>~happy_{~national}$$
**Null hypothesis**: people reading local human interest stories are **not** happier than people reading national human interest stories.

$$H_{0}:~~happy_{~local }~=~happy_{~national}$$

---

## Data

```{r}
## create data frame with happiness scores
d <- data.frame(
  local    = c(5, 6, 8, 4, 6, 9),
  national = c(2, 4, 3, 2, 3, 2)
)
```

| local| national|
|-----:|--------:|
|     5|        2|
|     6|        4|
|     8|        3|
|     4|        2|
|     6|        3|
|     9|        2|


---

## Calculating `t`

$$t = \frac{\bar{x}_{1} - \bar{x}_{2}}{\sqrt{
\frac{s_{1}^2}{n_{1}} +
\frac{s_{2}^2}{n_{2}}}}$$


```{r}
## group means (outermost parens added for printing)
(x1 <- mean(d$local))
(x2 <- mean(d$national))

## variances (s^2 = variance. s = sd)
(s1 <- var(d$local))
(s2 <- var(d$national))

## group lengths (should be same)
(n1 <- length(d$local))
(n2 <- length(d$national))
```

---

$$t = \frac{\bar{x}_{1} - \bar{x}_{2}}{\sqrt{
\frac{s_{1}^2}{n_{1}} +
\frac{s_{2}^2}{n_{2}}}}$$


```{r}
## calculate numerator
(num <- mean(x1) - mean(x2))

## calculate denominator
(denom <- sqrt((s1 / n1) + (s2 / n2)))

## t is equal to num divided by denom
(tstat <- num / denom)
```

---

## p-value(s)

Calculate the p-value for `tstat` using `pt()`

```{r}
## calculate degrees of freedom
(df <- n1 + n2 - 2)
```

Two-tailed

```{r}
## p-value for two-tailed (sided) test
2 * pt(-abs(tstat), df)
```

One-tailed (greater than)

```{r}
## p-value for one-tailed (sided) greater than test
pt(tstat, df, lower.tail = FALSE)
```

---

## `t.test()`

Compare with output from `t.test()`

```{r}
## t-test from stats pkg
t.test(d$local, d$national, alternative = "two.sided", var.equal = TRUE)
```

> *Note: We assume equal variances; this isn't always the case (see: `var.equal` documentation for `?t.test`)*


---

## Your turn

Compare the number of news website visits for Facebook versus Twitter users.

| fb_users| tw_users|
|--------:|--------:|
|       10|        3|
|        1|        8|
|        6|        1|
|        3|        0|
|        1|       10|
|        3|        7|

+ Open the [practice notebook](../assignment/02-practice.Rmd)

---
class: inverse, center, middle

## Chi-square tests

---

## Chi-square test

The t-test is only appropriate when the outcome/variable of interest is ordinal/interval/ratio.

To compare categorical responses between two groups, use **chi-square tests**.

+ Chi-square tests compare the expected frequencies (assuming uniform random distribution) with the observed frequencies (actually observed counts)

---

## Frequency matrix

In R, this is done by constructing a matrix where rows represent one variable dimensions and columns represent another variable's dimensions.

```{r}
## create frequency table
x <- matrix(c(12, 5, 2, 8, 6, 4, 10, 7), ncol = 2)

## row names
row.names(x) <- c("Democrat", "Lean Democrat", "Republican", "Lean Republican")

## column names
colnames(x) <- c("cnn", "fox")

## print matrix
x
```

---

## `chisq.test()`

Conducting a chi-square test is simple:

```{r}
## conduct chi-square test
chisq.test(x)
```

