---
title: "Day 4"
subtitle: "Correlations and factor analysis"
author: '`r rmdees::title_author()`'
date: '`r rmdees::title_icons()`'
output:
  xaringan::moon_reader:
    lib_dir: lib
    css: ["robot", "robot-fonts", "css/slides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = "#>")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
library(tidyverse)
```

## Agenda

---

## Agenda
1. Co-variance
   - What is covariance?
   - Calculating covariance
1. Correlations
   - Bivariate correlations
   - Correlation matrices
1. Factor analysis
   - What is a factor?
   - Calculating scores
   - Reliability
---
class: inverse, center, middle

## Co-variance

---

## Covariance

**Variance**: average distance from the mean

**Co-variance**: average


---

## Cov

$$cov(x, y) = \frac{\Sigma{(x-\bar{x}_{x})(y-\bar{y}_{y})}}{n - 1}$$

Where $\mu$ is actually the sample mean

---

## Cor

Correlation is the covariation divided by the total variation

$$cor(x, y) = \frac{cov(x, y)}{s_{x}s_{y}}$$

Which simplifies down to...

$$= \frac{{s_{xy}}}{\sqrt{s^2_{x}s^2_{y}}}$$

Where $s^2$ is variance and $s$ is standard deviation.

---

## Correlation

A **correlation** describes how variation in one variable relates to variation in another variable. 

In statistics, correlations describe the **magnitude** and **direction** of a relationship between two variables. 

---

```{r fig.height=5.5, fig.width=10}
## 20 observations from random normal distribution
x <- rnorm(20)
## x + 20 observations from random normal distribution
y <- rnorm(20) + x
## data frame
df <- data.frame(x, y)
## scatter plot
ggplot(df, aes(x, y)) + 
  geom_point() + 
  labs(title = "Scatter plot of y by x")
```

---

## Correlation coefficient: `r`

**`r`** (correlation coefficient): a statistic that describes the magnitude (size) and direction (order) of a relationship

There are several types of correlation coefficients, but all assume following:

+ The **range of possible values** goes from -1 to +1
+ **Direction** is described by the sign of the coefficient (negative = negative relationship and vice versa)
+ **Magnitude** is described by the coefficient's distance from zero
    + A correlation of zero indicates an absence of correlation (no relationship)

.footnote[In statistics, a **coefficient** refers to the multiplicative factor (a numeric estimate) for one variable in relation to another variable.]

---

## Types of correlations

**Pearson product-moment**: linear relationship between two variables 
$$x\sim\sim y$$

**Spearman's rho**: relationship between rankings of two variables 
$$x\sim\sim y$$

**Intraclass**: relationship between paired observations 
$$x_{t1}\sim\sim x_{t2}$$

---

## Semester

All of the correlations are similar, but you can assume when I talk about correlations I'm talking about the Pearson product-moment correlation.

When your variables are ordinal (ordered categories), it's important to remember there's a better way to estimate the correlation coefficient.

---

## Interval/ratio data

```{r}
## 20 observations from random normal distribution
x <- rnorm(20)
## x + 20 observations from random normal distribution
y <- rnorm(20) + x
## create data frame
df <- data.frame(x, y)

## print
head(df)
```



---

## Ordinal data

```{r}
## sample twenty values from numbers 1,2,3,4,5,6
x <- sample(c(1, 2, 3, 4, 5, 6), 20, replace = TRUE)
## sample twenty more, add them with x, and then divide by 2
y <- round((x + sample(c(1, 2, 3, 4, 5, 6), 20, replace = TRUE)) / 2, 0)
## create data frame
df_ordered <- data.frame(x, y)

## print
head(df_ordered)
```


---

## `cor(method = "pearson")`

By default, `cor()` returns the Pearson product-moment correlation.

```{r}
## using cor defaults
cor(df$x, df$y)
```

Use `method` to specify the type of correlation and `use` to deal with missing data

```{r}
## pearson product-moment correlation
cor(df$x, df$y, method = "pearson", use = "complete.obs")
```

---

## `cor(method = ` "spearman"`)`

Pearson product moment correlation coefficient.

```{r}
## pearson product-moment correlation
cor(df_ordered$x, df_ordered$y, method = "pearson")
```

And Spearman's rho (rank) correlation coefficient.

```{r}
## spearman rho correlation
cor(df_ordered$x, df_ordered$y, method = "spearman")
```


---

## Hypothesis testing

To conduct a signficiance test of a correlation using R, use `cor.test()`. 

```{r}
## correlation test
cor.test(df$x, df$y)
```

.footnote[To convert the output to a more usable data frame, use `tidy()` from the **{broom}** package.]

---

## `method =` "pearson"

Correlation test of interval/ratio data
```{r}
## correlation sig test
cor.test(df$x, df$y, method = "pearson") %>%
  broom::tidy() %>% head()
```

---

## `method =` "spearman"
 
Correlation test of ordinal data
```{r}
## correlation sig test
cor.test(df$x, df$y, method = "spearman") %>%
  broom::tidy() %>% head()
```


---

## Correlation tool

Visualize correlations of different values (-1 to 1):

+ https://mikewk.shinyapps.io/correlation/

---

## Practice

Guess the correlation coefficient (the game):

+ http://guessthecorrelation.com/




---
class: inverse, center, middle

## Factors and factor analysis

---

## Factor

A **factor** is a psychometric term for variable. 

In statistics, factors describe the **latent** variables which we are attempting to measure. 


---

## Items

An **item** refers to a single question/prompt/response-provoking stimulus in a questionnaire.

In the context of a study, a **variable** refers to a construct (or factor) to be examined.

Variables can consist of one or more items from a questionnaire.


---

## Multi-item variables

When a variable (or factor) is measured using multiple items, we still want to represent it with one number.

**How can we represent a variable measured with 5 likert-type items using one number per observation (respondent)?**


---

## Example

Our variable of interest is extraversion/introversion. We measure it using four likert-type items:

+ I like talking to lots of people
+ I like attending events where I meet knew people
+ I am comfortable at a party where I don't know anyone else
+ Meeting my friends' friends makes me nervous


---

## Example responses

```{r}
## first person's responses
person1 <- c(1, 2, 1, 7)

## second person's responses
person2 <- c(6, 6, 5, 2)
```


---

## Factor analysis

To make sure we have **reliable** measures, we use factor analysis.

**Factor analysis** essentially finds the correlation between responses for similar items. 

There are lots of details and variations, but knowing this much will help you in the future!

