---
title: "Day 5"
subtitle: "Data wrangling"
author: '`r rmdees::title_author()`'
date: '`r rmdees::title_icons_fa5()`'
output:
  xaringan::moon_reader:
    lib_dir: lib
    css: ["robot", "robot-fonts", "css/slides.css"]
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = "#>")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

## load libraries, set ggplot2 theme, and create datatable function
library(tidyverse)
theme_set(tfse::theme_mwk(base_size = 18))
set.seed(20180911)
print <- function(x, n = NULL, align = "c", digits = 3) {
  if (is.null(n)) n <- nrow(x)
  if (nrow(x) < n) n <- nrow(x)
  cat(paste(knitr::kable(x[seq_len(n), ], format = "markdown", 
    align = align, digits = digits), collapse = "\n"))
}
```


## Agenda
+ Computing basics
+ Tidyverse
   - Visualizing
   - Wrangling Basics
   - Merging & Joining
   - Transforming & Cleaning

---
class: inverse, center, middle

## Computing basics

---

## File/directory awareness
+ Every computer has a file/directory system
+ There's a default or home folder
   ```{r}
   ## use this to see your home folder
   normalizePath("~")
   ```
+ Files are organized like a tree
   - You can move from one branch to another
   - Moving from one folder to another is linear
+ Bash shortcuts
   - Use `..` to go back a folder or `../..` to go back two folders
   - Use `~/` as a shortcut for home

---

## Don't fall asleep yet!
Why do you need to know about your computer's file directory system?

+ Because you need to know where things are to interact with them!
+ Because you want to be able to replicate your work!

> **tl;dr**: open your file browser. See how they're organized? Yeah? Good!

---

## Don't let these terms confuse you
+ `File` is the **name**/location of a file
+ `Path` is the name/**location** of a file
+ `Folder` is the **name**/location of a file folder
+ `Directory` is the name/**location** of a file folder

<p class="note">`file == path & folder == directory`</p>

---
class: inverse, center, middle

## Do unto <huge>future you</huge> </br> as you would have future you </br> do unto <huge> you</huge>.

---

## FutuRe you

+ Using scripts and writing clear code will make your life easier
+ It's like writing down a routine of pointing and clicking that you don't have to memorize
+ The internet is full of routines/scripts that you can edit and customize to your liking
+ With some notes and some extra attention paid to make sure things are clear, future you will be very happy with current you!



<style>
huge {font-size:1.7em;}
</style>


<!--
## Data wrangling
+ Computing basics
+ Data frames
   - Wrangle: `select()`, `filter()`, `arrange()`, `mutate()`, `transmute()`
   - Bind: `bind_rows()`, `bind_cols()`
   - Join: `left_join()`, `full_join()`, `right_join()`
+ Text
   - Capitalize: `to_upper()`, `to_lower()`
   - RegEx: `str_replace()`, `str_remove()`, `str_extract()`, `str_detect()`
   - Split: `str_split()`
+ Numbers
   - Numbers (dollar sign, commas, etc.)
   - Dates/Times
-->