---
title: "Day 9"
subtitle: "Generalized linear models"
author: '`r rmdees::title_author()`'
date: '`r rmdees::title_icons_fa5()`'
output:
  xaringan::moon_reader:
    lib_dir: lib
    css: ["robot", "robot-fonts", "css/slides.css"]
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = "#>")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

## load libraries, set ggplot2 theme, and create datatable function
library(tidyverse)
theme_set(tfse::theme_mwk(base_size = 18))
set.seed(20180911)
print <- function(x, n = NULL, align = "c", digits = 3) {
  if (is.null(n)) n <- nrow(x)
  if (nrow(x) < n) n <- nrow(x)
  cat(paste(knitr::kable(x[seq_len(n), ], format = "markdown", 
    align = align, digits = digits), collapse = "\n"))
}
```


## Generalized linear models

---

## Agenda
+ Moderation (interactions)
+ Mediation
+ Generalized models
+ Logistic regression

---
class: inverse, center, middle

## Moderation

“Moderation analysis” used to examine how the
effect of an antecedent
X on an consequent
Y depends
on a third moderator variable
M (a.k.a. “interaction”)

---

## What is moderation?

Moderation. The effect of X on Y can be said to be moderated if its size or
direction is dependent on M. It tells us about the conditions that facilitate,
enhance, or inhibit the effect, or for whom or what the effect is large vs. small,
present versus absent, and so forth

```{r}
foo("Newsworthiness", data = "data", 
  c("Twitter Use")) + 
  ggsave("static/img/x-y.png", width = 8, height = 4, units = "in")

foo("Newsworthiness", data = "bias", 
  c("Twitter Use", "Journalist")) + 
  ggsave("static/img/x12-y.png", width = 8, height = 4, units = "in")

foo("Newsworthiness", data = "bias", 
  c("Twitter Use", "Journalist", "News Use")) + 
  ggsave("static/img/x13-y.png", width = 8, height = 4, units = "in")
```

M is depicted here to moderate the size of the effect of X on Y, meaning that the size of
the effect of X on Y depends on M. We say M is the moderator of the X → Y
relationship, or X and M interact in their influence on Y.

---
class: inverse, center, middle

## Mediation


A mediation model links a putative cause (X) to a presumed
effect (Y) at least in part through an intermediary or “mediator”
variable (M)


X→M→Y is a causal chain of events. A mediator variable can be a psychological
state, a cognitive process, an affective response, a biological change, or any other
conceivable “mechanism” variable through which X exerts an effect on Y. But it
must be causally between X and Y.


The indirect effect quantifies the effect of X on Y through M. Evidence that ab is different
from zero is consistent with mediation. Evidence that path c is different from zero is not a
requirement of 21st century mediation analysis. Correlation between X and Y is neither
sufficient nor necessary to claim that X affects Y.

---
class: inverse, center, middle

## Generalized models

---



