---
title: "Group comparisons"
output: html_notebook
---

The data set `nwv`, entered below, provides the number of news website visits logged for six Facebook users and six different Twitter users.

```{r}
## news website visits (counts) of facebook and twitter users
nwv <- data.frame(
  fb_users = c(10, 1, 6, 3, 1, 3),
  tw_users = c(3, 8, 1, 0, 10, 7)
)
```

1. Calculate (and store) the mean `x1` `x2`, variance `s1` `s2`, and size `n1` `n2` of each group.

```{r}
## I started this for you; complete the rest of the code

## group means
x1 <- 
x2 <- 

## group variances
s1 <- 
s2 <- 

## group sizes
n1 <- 
n2 <- 
```

2. Using the objects created for the previous question, calculate (and store as `num` and `denom`) the numerator and denominator based on the independent samples t-test formula

```{r}
## your code goes here
```

3. Using `num` and `denom`, calculate (and store as `tstat`) the t statistic

```{r}
## your code goes here
```

4. Calculate (and store as `df`) the degrees of freedom (you can assume equal variances)

```{r}
## your code goes here
```

5. Use `tstat` and `df` to calculate the p-value for a two-sided hypothesis

```{r}
## your code goes here
```

6. Visualize the means or distributions of the two groups. Make sure the values associated with each group are clearly marked.

```{r}
## load tidyverse
library(tidyverse)

## gather data long-wise and then construct boxplot
nwv %>%
  gather(group, visits) %>%
  ggplot(aes(x = group, y = visits, fill = group)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(x = "User group", y = "Number of news website visits", 
    title = "Comparing news website visits for Twitter versus Facebook users", 
    subtitle = "Estimates based on browser histories of twelve total users") + 
  theme(plot.title = element_text(face = "bold"), 
    legend.position = "none")
```