---
title: "R Notebook"
output: html_notebook
---

## the data


```{r}
if (!requireNamespace("broom", quietly = TRUE)) {
  install.packages("broom")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
## load tidyverse
library(tidyverse)

## read in data
dat <- readRDS("https://github.com/mkearney/stat/raw/master/static/data/class-data.rds")
```


## 1. Visualize partisan score by party group

```{r}
pp_levels <- c(
  "Strong Democrat",
  "Democrat",
  "Lean Democrat",
  "Slight Lean Democrat",
  "Slight Lean Republican",
  "Lean Republican",
  "Republican",
  "Strong Republican"
)

dat %>% 
  mutate(party_label = factor(party_label, levels = pp_levels)) %>%
  ggplot() +
  aes(x = party_label, y = therm_1 - therm_2) +
  geom_boxplot() + 
  coord_flip()
```

## 2. Conduct ANOVA model predicting partisan by partisan (therm_1 - therm_2)

```{r}
dat$partisan <- dat$therm_1 - dat$therm_2

m1 <- aov(partisan ~ party_label, data = dat)
```

## 3. View ANOVA table

```{r}
anova(m1)
```

## 4. View summary

```{r}
summary(m2)
```


## 5. Visualize `ambiv_sexism` by `party_label`

```{r}

```


## 6. Conduct ANOVA of `ambiv_sexism` by `party_label`

```{r}

```

## 7. View ANOVA table

```{r}

```

## 8. View summary

```{r}

```



## 9. Conduct ANOVA of `polar_therm` by `sex` and by `party_label`

```{r}

```

## 10. View ANOVA table

```{r}

```

## 11. View summary

```{r}

```




## 12. Conduct ANOVA of `polar_therm` by interaction of `sex` and `party_label`

```{r}

```

## 13. View ANOVA table

```{r}

```

## 14. View summary

```{r}

```


